\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}

\title{Machine Learning in Space Weather}
\author{Mandar Chandorkar}
\date{January 21 2019}

\begin{document}

\maketitle

The study of variations in the space environment between the Sun and the Earth constitutes the core of \textit{space weather} research. Ionized plasma ejected by the Sun couples with the Earth’s magnetic field in complex processes that determine the state of the Earth's magnetosphere. Adverse effects from space weather can impact communication networks, power grids and logistics infrastructure, all crucial pillars of a civilization that is reliant on technology.

It is thus important to leverage data sources, scientific knowledge and statistical learning methodology to create forecasting and monitoring systems for the upcoming space age. This thesis aims to be a step towards that goal. The work is organised into the following sections/chapters.

\begin{enumerate}
\item \textit{Forecasting}: We develop probabilistic forecasting models for predicting \textit{geo-magnetic} time series. Combining ground based and satellite measurements, we propose a \textit{gaussian process} model for one hour ahead prediction of the \textit{Dst} time series \cite{ChandorkarDst}, \cite{CHANDORKAR2018237}. We augment this model with a \textit{long short-term memory} network and produce probabilistic predictions upto six hours ahead for \textit{Dst} \cite{doi:10.1029/2018SW001898}.

\item \textit{Parameter Inference \& Uncertainty Quantification}: Quantifying uncertainties in the Earth's \textit{radiation belt} parameters is an important step for producing ensembles of high fidelity simulations of the \textit{plasmasphere}. We combine simplified dynamical models with markov chain monte carlo techniques to infer uncertainties in plasmasphere parameters, using data from probes orbiting in the radiation belts.

\item \textit{Causal Time Lag Prediction}: In temporal phenomena, it is often the case that causal effects of events are not immediately observed, but after a certain time interval which can be dynamic. One prominent example of such behavior is the \textit{Sun-Earth} system. Particles ejected from the Sun, also called the \textit{solar wind}, reach the Earth's plasmasphere after a time delay which is uncertain. We propose a novel neural network based method, for predicting causal time delay between time series and apply it to the problem of \textit{solar wind} propagation.



\end{enumerate}




\bibliographystyle{plain}
\bibliography{M335}

\end{document}
